---
title: "HW-1"
author: "Ruhani Arora"
date: '2022-09-28'
output: github_document
---

```{r}
library(dplyr)
library(nycflights13)
```


1.How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
flights %>% 
  filter(is.na(dep_time))
```
The other missing variables are dep_delay, arr_time, arr_delay and air_time. These rows might represent that the flight was canceled which is why these variables are missing.


2.Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flights %>% 
  mutate(
    dep_time_mins= (dep_time%/% 100 * 60) + dep_time %% 100,
    sched_dep_time_mins = (sched_dep_time%/% 100 * 60) + sched_dep_time %% 100
    )
```


3.Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

```{r}
library(ggplot2)
cancelled_flights_per_day <- flights %>%
  mutate(cancelled_flight = (is.na(dep_delay) | is.na(arr_delay))) %>%
  group_by(year,month,day) %>%
  summarise(
    cancelled_flights = sum(cancelled_flight),
    total_flights = n(),
  ) 
ggplot(cancelled_flights_per_day)+
  geom_point(aes(x=total_flights,y=cancelled_flights))
```

After looking at the graph, there seems to be a pattern between the days having more number of flights to have a higher number of cancelled flights.

```{r}
prop <- flights %>%
  mutate(cancelled_flight = (is.na(dep_delay) | is.na(arr_delay))) %>%
  group_by(year,month,day) %>%
  summarise(
    cancelled_prop = mean(cancelled_flight),
    avg_dep_delay = mean(dep_delay, na.rm = TRUE)
  ) %>%
  ungroup()
ggplot(prop) +
  geom_point(aes(x = avg_dep_delay, y = cancelled_prop))
```

The proportion of canceled flights related to the average dep delay as if there is a longer delay in departure time the flight is more likely to get canceled. However, there are a few outliers in this graph where the avg dep delay is not high but the cancellation prop is is highest.